---
output:
  pdf_document: default
  html_document: default
---


```markdown
---
title: "Data Science Project: The Scent of Data"
author: "Arwa, Marh, Layan"
date: "2024-12-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load required libraries
library(tidyverse)
library(janitor)
library(ggplot2)
library(naniar)
library(broom)
library(patchwork)

# Ensure reproducibility
set.seed(123)

# Load and clean dataset
parfumo_data_clean <- read_csv("parfumo_data_clean.csv") %>%
  clean_names()

# Handle missing values (basic cleaning)
parfumo_data_clean <- parfumo_data_clean %>%
  mutate(
    release_year = as.integer(release_year),  # Convert to integer for modeling
    rating_value = as.numeric(rating_value),  # Ensure numeric type
    rating_count = as.numeric(rating_count),  # Ensure numeric type
    main_accords = replace_na(main_accords, "Unknown"),  # Replace missing accords
    top_notes = replace_na(top_notes, "Unknown"),
    middle_notes = replace_na(middle_notes, "Unknown"),
    base_notes = replace_na(base_notes, "Unknown")
  )


# Quick inspection of the dataset
glimpse(parfumo_data_clean)

```

# Introduction

In this report, we analyze the Parfumo Fragrance Dataset to uncover patterns in perfume ratings and composition. Our primary goal is to understand the factors that influence user ratings and build a predictive model to estimate perfume ratings based on available attributes. This report is structured as follows:

1. **Problem Statement and Background: Defining the problem and referencing related work.
2. **Data**: Description of the dataset, variables, and potential issues.
3. **Analysis**: Methods and tools employed to clean, explore, and model the data.
4. **Results**: Findings presented with visualizations and tables.
5. **Discussion**: Conclusions, limitations, and potential next steps.

---

# Problem Statement and Background

Perfume ratings are subjective and influenced by various factors, including scent composition, release year, and user preferences. Understanding these factors can help brands tailor their products to consumer preferences.

**Problem Statement**:  
We aim to identify key drivers of perfume ratings and build a model to predict ratings based on attributes such as scent composition, perfumers, and release year.

**Related Work**:
1. Prior studies have explored the relationship between olfactory notes and user preferences, emphasizing the importance of top and base notes in lasting impressions.
2. Predictive modeling in other domains (e.g., movies, books) demonstrates the effectiveness of structured metadata for reliable predictions.

---

# Data

## Dataset Overview
The dataset contains 13 variables and over 5,000 observations. Each row represents a perfume, with metadata including its name, brand, release year, scent composition (top, middle, and base notes), user ratings, and the number of ratings.

## Key Variables
- **Outcome Variable**:  
  `Rating_Value` (continuous numeric)  
  - Measured on a 5-point scale.
  - Represents the average user rating.
  - Source: Parfumo user reviews.

- **Predictor Variables**:  
  - `Release_Year` (numeric): Year of release.  
  - `Rating_Count` (numeric): Number of user ratings.  
  - `Main_Accords`, `Top_Notes`, `Middle_Notes`, `Base_Notes` (categorical): Scent components.  
  - `Brand` and `Perfumers` (categorical): Metadata about the perfume's creators.  

---


## Summary of Predictor Variables
```

# Top 10 Main Accords
parfumo_data_clean %>%
  filter(main_accords != "Unknown") %>%  # Exclude missing accords
  count(main_accords, sort = TRUE) %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(x = reorder(main_accords, n), y = n)) +
  geom_col(fill = "orchid") +
  coord_flip() +
  labs(
    title = "Top 10 Main Accords",
    x = "Main Accord",
    y = "Count"
  )
```

---

# Data Issues

### Missing Data
- Missing values are present in `Top_Notes`, `Middle_Notes`, and `Base_Notes`.  
  **Mitigation**: Removed or imputed missing data for analyses.

### Bias
- User ratings may be influenced by brand reputation or availability.

### Imbalanced Features
- Some brands and perfumers are overrepresented.  
  **Mitigation**: Normalization and balanced sampling where required.

---

# Analysis

## Methods and Tools
We used the following methods:
1. **Exploratory Data Analysis (EDA)**:
   - Visualized distributions and relationships between variables.
2. **Linear Regression**:
   - Built a predictive model for `Rating_Value`.
3. **Feature Importance**:
   - Used correlation and regression coefficients to identify key predictors.

---

## Missing Values
```{r missing-values-analysis}
# Visualize missing data
gg_miss_var(parfumo_data_clean) +
  theme_minimal() +
  labs(
    title = "Missing Data Overview",
    x = "Variables",
    y = "Number of Missing Values"
  )

```

---

# Advanced Analysis

## Summary of Outcome Variable
```{r outcome-summary}
# Outcome variable distribution
parfumo_data_clean %>%
  filter(!is.na(rating_value)) %>%
  ggplot(aes(x = rating_value)) +
  geom_histogram(binwidth = 0.1, fill = "steelblue", color = "white") +
  theme_minimal() +
  labs(
    title = "Distribution of Perfume Ratings",
    x = "Rating Value",
    y = "Frequency"
  )

```

## Summary of Predictor Variables
```{r predictor-summary}
# Distribution of predictor variables
# Release Year
parfumo_data_clean %>%
  filter(!is.na(release_year)) %>%
  ggplot(aes(x = release_year)) +
  geom_histogram(binwidth = 5, fill = "orchid", color = "white") +
  theme_minimal() +
  labs(
    title = "Distribution of Perfume Release Years",
    x = "Release Year",
    y = "Frequency"
  )

```

---

# Predicting Ratings Using Multiple Features
```{r multiple-regression}
# Multiple regression model
multi_lm_model <- lm(
  rating_value ~ rating_count + release_year + as.factor(main_accords),
  data = parfumo_data_clean %>%
    filter(!is.na(rating_value))  # Exclude missing values
)
summary(multi_lm_model)

# Visualization of predicted vs. actual ratings
parfumo_data_clean <- parfumo_data_clean %>%
  mutate(predicted_rating = predict(multi_lm_model, newdata = .))

ggplot(parfumo_data_clean, aes(x = rating_value, y = predicted_rating)) +
  geom_point(alpha = 0.5, color = "steelblue") +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Actual vs Predicted Ratings",
    x = "Actual Ratings",
    y = "Predicted Ratings"
  )

```

---

# Discussion

### Conclusions
- Perfume ratings are strongly influenced by the number of reviews.
- Certain scent families consistently receive higher ratings.

### Limitations
- Data is sourced from a single platform, limiting generalizability.
- Missing data in olfactory notes may obscure the influence of specific components.

### Next Steps
1. Incorporate user demographics or review text for deeper insights.
2. Develop advanced predictive models, such as Random Forest or Gradient Boosting.
3. Implement a `shiny` app for interactive exploration of perfumes.

```

---
