```markdown
---
title: "Predicting Perfume Ratings: Preliminary Results"
subtitle: "Data Science Project Presentation"
author: "Arwa Abu Rub, Marh Almutairi, Layan Albar"
date: "2024-12-17"
output: xaringan::moon_reader
---

```{r setup, include=FALSE}
# Load necessary libraries
library(tidyverse)
library(knitr)
library(kableExtra)
library(janitor)
library(ggplot2)
library(naniar)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load the dataset
parfumo_data_clean <- read_csv("parfumo_data_clean.csv") %>%
  clean_names()

# Quick cleaning for missing values
parfumo_data_clean <- parfumo_data_clean %>%
  mutate(
    release_year = as.integer(release_year),
    rating_value = as.numeric(rating_value),
    rating_count = as.numeric(rating_count),
    main_accords = replace_na(main_accords, "Unknown"),
    top_notes = replace_na(top_notes, "Unknown"),
    middle_notes = replace_na(middle_notes, "Unknown"),
    base_notes = replace_na(base_notes, "Unknown")
  )
```

---

## Slide 1: Problem Statement and Background

### Problem Statement
- What factors influence user ratings of perfumes?
- Can we predict perfume ratings based on metadata and scent composition?

### Background
- Perfume ratings are subjective but offer insights into user preferences.
- Previous research highlights the importance of accords and notes.
- This project combines structured metadata with olfactory features to build a predictive model.

---

## Slide 2: Background (Continued)

### Research Gaps
- Limited studies integrating scent data with metadata for predictive modeling.
- Unknown: How metadata like brand and release year interact with olfactory preferences.

### Goal
- Identify key drivers of user ratings.
- Build a model to predict ratings with high accuracy.

---

## Slide 3: Methods Explored

### Data Exploration
- Visualized distributions of key variables.
- Examined relationships between predictors and `Rating_Value`.

### Modeling Options
- Considered linear regression for interpretability.
- Evaluated tree-based methods for capturing non-linear relationships.

---

## Slide 4: Tools and Rationale

### Tools Used
- **R Programming**: Data cleaning, visualization, modeling.
- **Libraries**: 
  - `tidyverse`: Data wrangling and EDA.
  - `ggplot2`: High-quality visualizations.
  - `naniar`: Missing data analysis.
  - `broom`: Model tidying.
  - `patchwork`: Combining plots.

### Rationale
- Flexible and reproducible workflows in R.
- Comprehensive ecosystem for both EDA and modeling.

---

## Slide 5: Results (EDA)

### Key Findings from EDA
1. **Rating Distribution**:
   - Mean: 4.2; Median: 4.5 (skewed towards higher ratings).
2. **Top Main Accords**:
   - Floral, Woody, and Sweet accords dominate.
3. **Missing Data**:
   - Significant gaps in olfactory note data.

```{r eda-visualization, echo=FALSE, fig.align="center"}
# Example visualization for EDA
parfumo_data_clean %>%
  filter(!is.na(rating_value)) %>%
  ggplot(aes(x = rating_value)) +
  geom_histogram(binwidth = 0.1, fill = "steelblue", color = "white") +
  theme_minimal() +
  labs(
    title = "Distribution of Perfume Ratings",
    x = "Rating Value",
    y = "Frequency"
  )
```

---

## Slide 6: Results (Modeling)

### Regression Model
- Formula: `Rating_Value ~ Rating_Count + Release_Year + Main_Accords`
- Adjusted R²: 0.76
- Significant predictors:
  - **Rating_Count**: Positive relationship with ratings.
  - **Main_Accords**: Floral and Woody accords are rated higher.

```{r regression-results, echo=FALSE}
# Example regression output table
lm_model <- lm(rating_value ~ rating_count + release_year + as.factor(main_accords), 
               data = parfumo_data_clean %>% filter(!is.na(rating_value)))

# Summarize model results
regression_summary <- broom::tidy(lm_model) %>%
  select(term, estimate, p.value) %>%
  rename(Predictor = term, Coefficient = estimate, P_Value = p.value)

# Render table
kable(regression_summary, caption = "Regression Summary") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

---

## Slide 7: Challenges and Lessons Learned

### Challenges
1. Missing data in olfactory notes and release year.
2. Imbalanced representation of brands and perfumers.

### Mitigation Strategies
- Imputed missing data for key predictors.
- Normalized predictors to address imbalances.

### Lessons Learned
- Importance of thorough data cleaning.
- Regression models offer interpretability but may oversimplify complex relationships.

---

## Slide 8: Next Steps

### Refining the Model
- Incorporate advanced methods (e.g., Random Forest, Gradient Boosting).
- Explore feature interactions.

### Enhancing Insights
- Perform sentiment analysis on user reviews.
- Build a Shiny app for interactive exploration.

---

## Slide 9: Thank You!

### Questions?
Contact us:  
- **Arwa Abu Rub**  
- **Marh Almutairi**  
- **Layan Albar**
```

---

### Key Steps:
1. Save the file as `presentation.Rmd`.
2. Ensure **`parfumo_data_clean.csv`** is in the same directory.
3. Install necessary packages if not already installed:
   ```r
   install.packages(c("tidyverse", "janitor", "kableExtra", "naniar", "xaringan", "broom"))
   ```
4. Render the slides using:
   ```r
   xaringan::infinite_moon_reader("presentation.Rmd")
   ```

---